Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Research is supported by Russian Foundation of Basic Research, grant No 14-01-31449~mol\_a.}

### 2015-09-06 paper.tex
Removed last questions.

% first the title is needed
\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{\textbf{Ok .....................!!!!!!!!!!!!!!!!!!..................}}}
% a short form should be given in case it is too long for the running head
\titlerunning{ATP Based on the PCF Calculus}

@@

\textbf{Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics. But type quantifiers are stable in languages of another applied fields.}

@@

If a PCF contains a question with disjunctive branching, then its base subformula will be divided into several new base subformulas during searching procedure of the logical inference. The number of new subformulas corresponds to the number of disjunctive subformulas in the consequent of the question. In a straightforward variant of the prover implementation, the sharing will require copying previous state of the original formula more than once; this copying have a linear complexity, and it leads to the redundant expenses of memory and processor time. PST allows us to implement common subbranches sharing instead of copying. Let's consider the following base subformula:
$$\exists: A(a) - \forall x: A(x) - \left\{
\begin{array}{lcl}
 \exists \colon B(x) & - & \forall y: B(y) - \exists\colon\boldsymbol{False}\\
 \exists \colon C(x) & - & \forall y: C(y) - \exists\colon\boldsymbol{False}.
\end{array}
\right. $$
The formula has disjunctive branching and it is deep. PST for the refutation of this base subformula is presented in Fig.~\ref{fig:datasharing2}.
\begin{figure}[h]
  \vspace{0.5cm}
  \centering
  \includegraphics[width=0.82\linewidth]{img/DataSharing2p1.eps}
  \caption{Example of PST for a base subformula refutation}
  \label{fig:datasharing2}
\end{figure}

@@

%=======================================================================
%==========================EXPERIMENTS==================================
%=======================================================================
\section{Experiments}
Testing the prover under development is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org). This library is the de facto standard for testing ATP provers. The type of problems that were chosen are the problems of FOF without equality. Total number of problems is 1221 with rating from 0.0 (very easy) to 1.0 (unresolved). Total number of solved problems is 780. The maximal rating of solved problem is 0.62. The best results among domains presented in Table~\ref{tbl:1}.

@@

%=============================
\subsubsection{Lazy concretization.}
Lazy concretization strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum. The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy. The decrease of the number of steps leads to a decrease of the problem solving time. \textbf{Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching}.

@@

Experiments were carried out on the logical problems which PCF--for\-ma\-li\-za\-ti\-ons having necessary properties for the parallel strategies utilization: disjunctive branching, large number of questions, large conjuncts in questions. In Fig.~\ref{fig:parallel} the results of experiments are shown. \textbf{In the figure ``Strategy 1'' denotes ... and ``Strategy 2'' denotes ...}

@@

%=======================================================================
%==========================CONCLUSION===================================
%=======================================================================
\section*{Conclusion}
In this paper, we presented the results of the positively constructed formulas (PCF) calculus investigation in the context of the automatic theorem proving. The main result is a prover program implemented based on the PCF calculus. Several logical inference strategies were devised, adopted from other papers and implemented in this prover: memory sharing, lazy concretization, $k,m$--condition, parallel strategies. A special global data structure (proof state tree) has been developed. Most of the strategies are implemented on the base of the structure.

@@

\bibitem{DPL2} Alexandrescu, A. The D Programming Language. -- Addison-Wesley Professional, 2010. -- 460p.

\bibitem{DPL1} Armstrong, J. D Programming Language. URL: http://www.digitalmars.com/d/.

\bibitem{Erlang1} Armstrong, J. Programming Erlang.  -- The Pragmatic Programmers, 2007. -- 519p.

\bibitem{TPTPTrans} Gelder, A.V. TPTPparser utility. 2006. URL: http://users.soe.ucsc.edu/\~{}avg/TPTPparser/.

\bibitem{gmemory} GLIB: Memory Slices. URL: http://developer.gnome.org/glib/2.34/glib-Memory-Slices.html.

\bibitem{subtree} Graf, P. Substitution Tree Indexing // Proceedings of the 6th International Conference on Rewriting Techniques and Applications. 1995. pp.~117--131.

@@

\bibitem{tptp} Sutcliffe, G. The TPTP Problem Library and Associated Infrastructure: The FOF and CNF Parts, v3.5.0 // Journal of Automated Reasoning. V43, N4, pp.337--362.

\bibitem{SNV1990} Vassilyev, S.N.: Machine Synthesis of Mathematical Theorems. The Journal of Logic programming, V.9, No.2--3, pp. 235--266 (1990)

\bibitem{ICDS2000} Vassilyev, S.N., Zherlov, A.K., Fedunov, E.A.,
Fedosov, B.E.: Intelligent Control of Dynamic Systems. Moscow: Fizmatlit (2000)(in Russian)

>>>>

% first the title is needed
\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Research is supported by Russian Foundation of Basic Research, grant No 14-01-31449~mol\_a.}}
% a short form should be given in case it is too long for the running head
\titlerunning{ATP Based on the PCF Calculus}

@@

Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics. But type quantifiers are stable in languages of another applied fields.

@@

If a PCF contains a question with disjunctive branching, then its base subformula will be divided into several new base subformulas during searching procedure of the logical inference. The number of new subformulas corresponds to the number of disjunctive subformulas in the consequent of the question. In a straightforward variant of the prover implementation, the sharing will require copying previous state of the original formula more than once; this copying have a linear complexity, and it leads to the redundant expenses of memory and processor time. PST allows us to implement common subbranches sharing instead of copying. Let's consider the following base subformula:
$$\exists: A(a) - \forall x: A(x) - \left\{
\begin{array}{lcl}
 \exists \colon B(x) & - & \forall y: B(y) - \exists\colon\boldsymbol{False}\\
 \exists \colon C(x) & - & \forall y: C(y) - \exists\colon\boldsymbol{False}.
\end{array}
\right. $$
The formula has disjunctive branching and it is deep. PST for the refutation of this base subformula is presented in Fig.~\ref{fig:datasharing2}.
\begin{figure}[h]
  \vspace{0.5cm}
  \centering
  \includegraphics[width=0.82\linewidth]{img/DataSharing2}
  \caption{Example of PST for a base subformula refutation}
  \label{fig:datasharing2}
\end{figure}

@@

%=======================================================================
%==========================EXPERIMENTS==================================
%=======================================================================
\section{Testing the prover}
Testing the prover under development is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org). This library is the de facto standard for testing ATP provers. The type of problems that were chosen are the problems of FOF without equality. Total number of problems is 1221 with rating from 0.0 (very easy) to 1.0 (unresolved). Total number of solved problems is 780. The maximal rating of solved problem is 0.62. The best results among domains presented in Table~\ref{tbl:1}.

@@

%=============================
\subsubsection{Lazy concretization.}
Lazy concretization strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum. The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy. The decrease of the number of steps leads to a decrease of the problem solving time. Application of the UHE constraint~1 (limiting depth of instantiation) is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 (storing UHE expression) in case if the formula contains no questions without disjunctive branching.

@@

Experiments were carried out on the logical problems which PCF--for\-ma\-li\-za\-ti\-ons having necessary properties for the parallel strategies utilization: disjunctive branching, large number of questions, large conjuncts in questions. In Fig.~\ref{fig:parallel} results of experiments are shown. In the figure ``Strategy 1'' denotes paralleling base subformula refutation and ``Strategy 2'' references to parallel answer search to a question.

@@

%=======================================================================
%==========================CONCLUSION===================================
%=======================================================================
\section*{Conclusion}
In this paper, we briefly presented the results of the positively constructed formulas (PCF) calculus investigation in the context of the automatic theorem proving. The main result is a prover program implemented based on the PCF calculus. Several logical inference strategies were devised, adopted from other papers and implemented in this prover: memory sharing, lazy concretization, $k,m$--condition, parallel strategies. A special global data structure (proof state tree) has been developed. Most of the strategies are implemented on the base of the structure.

@@

\bibitem{SNV1990} Vassilyev, S.N.: Machine Synthesis of Mathematical Theorems. The Journal of Logic programming, V.9, No.2--3, pp. 235--266 (1990)
\bibitem{ICDS2000} Vassilyev, S.N., Zherlov, A.K., Fedunov, E.A.,
Fedosov, B.E.: Intelligent Control of Dynamic Systems. Moscow: Fizmatlit (2000)(in Russian)
\bibitem{Bourbaki} Bourbaki, N.: Theory of Sets. Paris, Hermann (1968)
\bibitem{DPL1} Armstrong, J. D Programming Language. URL: http://www.digitalmars.com/d/.
\bibitem{DPL2} Alexandrescu, A. The D Programming Language. -- Addison-Wesley Professional, 2010. -- 460p.
\bibitem{Erlang1} Armstrong, J. Programming Erlang.  -- The Pragmatic Programmers, 2007. -- 519p.
\bibitem{gmemory} GLIB: Memory Slices. URL: http://developer.gnome.org/glib/2.34/glib-Memory-Slices.html.
\bibitem{tptp} Sutcliffe, G. The TPTP Problem Library and Associated Infrastructure: The FOF and CNF Parts, v3.5.0 // Journal of Automated Reasoning. V43, N4, pp.337--362.
\bibitem{TPTPTrans} Gelder, A.V. TPTPparser utility. 2006. URL: http://users.soe.ucsc.edu/\~{}avg/TPTPparser/.
\bibitem{subtree} Graf, P. Substitution Tree Indexing // Proceedings of the 6th International Conference on Rewriting Techniques and Applications. 1995. pp.~117--131.

@@
